A quantitative framework for investigating the reliability of empirical network construction

### 2018-11-28 manuscript/reply_to_reviewers.tex

	4. Demand for extension to weighted networks [[not done]]

		\begin{refquote}
		L45 It is a shame not to give some idea of how the method could be extended (or is this the intention of the authors?). For empirical analysis of networks it is rare to use binary networks given the value of weighted networks for taking account of sampling biases. Indeed, I am much less concerned about whether an interaction never occurs, and much more concerned with the frequency of occurrence.
		\end{refquote}

		\textbf{R:} As stated above, we do not agree that weighted networks reduce sampling biases. Moreover, we disagree that it is rare to use binary networks as most food webs (for example) are still binary. It takes an even greater sampling effort to quantify interaction strengths than the presence and absence of links, so for species-rich systems in particular there simply are not many high-quality weighted networks published. If the Reviewer is only concerned about weighted interactions, we congratulate them on their tractable study system. Nevertheless, we do not wish to exclude the many researchers who are not so fortunate.
		\smallskip
		To model frequencies of interactions, ...
			 [[Do we want to add a bit of non-binary stuff or just make it clearer that we are talking about feasibility and that probabilities of occurrance could be layered on later?]]

	5. Description of causes of uncertainty poorly integrated with the case study [[done]]

		\begin{refquote}
		L50 This is a nice and clearly explained description of the problem that many network ecologists ignore, or are unaware of.
		Section from L50. This section is valuable, but had very weak links with the empirical analysis later in the paper. For instance in the example of the gall-formers, there was not formal discussion about the different sources of uncertainty. 
		\end{refquote}

		\textbf{R:} As we state in the section \textbf{Estimating uncertainty}, we usually cannot distinguish different sources of uncertainty in empirical data. Those studies which attempt to do so (e.g.,~\citet{Graham2018}) remove process uncertainty by assuming constant interaction probabilities during a short, intense sampling period and also assume that we know the true set of feasible interactions. This may be the case for Graham and Weinstein's relatively small and well-studied system but is not likely to be true for systems containing hundreds of species, systems containing recently-arrived species, systems just beginning to be studied, etc.
		\smallskip
		To make this more clear, we have rephrased this section and added a line (lines XX-XX) stating that we usually cannot determine the cause of all the zeros in an interaction matrix (we can identify obviously unfeasible interactions, but likely cannot say whether a zero for a rare species is because we did not observe enough individuals, there were not enough individuals to perform a particular interaction, or because the interaction is acutally unfeasible). We hope that this revision makes it clearer that the description of different sources of uncertainty is a conceptual guide only.

		\begin{quotation}
			An observed link definitely occurred, but there are multiple reasons why a given link may not be observed \emph{whether or not the interaction truly occurred}. Moreover, given an unobserved interaction in an empirical dataset, we often cannot determine why the interaction did not occur \emph{post hoc}. The detection of any interaction is a stochastic process subject to many levels of uncertainty. As a conceptual guide to the factors affecting this process, we describe three nested levels of uncertainty that roughly address the questions: "Could species $i$ and $j$ interact?", "Do they interact?" and "Do we observe the interaction?" (Fig.~\ref{conceptual_fig}). 
		\end{quotation}

@@

			\end{quotation}

	8. Confusion about goal of modelling [[done]]

		\begin{refquote}
		I was struggling to really understand what the authors were really trying to model (partly because of the lack of a link between the theoretical and empirical parts of the paper). Are the authors trying to estimate L (the probability that an interaction is feasible or not – I would describe this as a ‘master’ or meta- network), or is it that they are trying to estimate X|L (the probability that the interaction locally occurs)? I would be interested in both, but the second aspect seemed to be ignored in the gall-former example.
		\end{refquote}

		\textbf{R:} In the case study, we are modelling L (the probability that an interaction is feasible). We do not have quantitative data within sites in our dataset, so we cannot isolate the probability of detecting an interaction at a site from the probability that the interaction occurs. This is also the probability we were most interested in. We do note, however, that one could also model X|L using the same framework if there were independent counts of co-occurrences and interactions within a single site. We now state throughout the manuscript that we are interested in compiling the metaweb, the set of interaction which can feasibly occur. In particular, this is now specified in the introduction. We hope that our goal is now crystal clear.

		\begin{quotation}
			To supplement repeated sampling efforts and reduce uncertainty around unobserved interactions, we suggest researchers use prior knowledge of the same or similar systems. Here, we outline a Bayesian approach to combining repeated sampling and prior knowledge in order to assemble a metaweb. We then demonstrate this approach using an intensively and repeatedly sampled empirical dataset. Finally, we show how a probabilistic understanding of a metaweb allows us to establish confidence intervals around measures of network structure.
			\end{quotation}

	9. Misunderstanding of how biased sampling does not solve problems of uncertainty [[done]]

@@

		\begin{refquote}
		L136 This is a completely different issue, that you do not address in the paper.
		\end{refquote}

		\textbf{R:} We do not believe that biased sampling is a completely different issue as the tendency for some interactions to be sampled more easily than others clearly contributes to varying levels of detection uncertainty. Increasing sampling effort will not do much good if we repeatedly sample the same common interactions and very rarely identify any new interactions. We now address questions about whether or not sampling more will remove uncertainty in a dedicated box. We still mention the fact that increased sampling will not reduce uncertainty evenly across interactions as this is an important point to consider when designing sampling protocols and when analysing empirical data. We also point out that combining sampling types is a partial solution to this problem; if different sampling methods have different biases, they may partially cancel each other out. We believe that this addresses the question as much as possible given our focus on analysing collected data rather than collecting new data.

@@

		\textbf{R:} We have expanded our description of different types of priors and when researchers might prefer one over another. Following a comment from Reviewer 1 we have removed much of the mathematical background for the Bayesian approach, but this material is still present in the supplemental material. We hope that this simplifies the approach adequately.

	11. Integrate L155ff with previous section [[done]]

		\begin{refquote}
		L155ff I think it would be helpful for this section to be integrated with the previous section. It relies on the reader putting quite a lot of work in to understand the link between the two.
		\end{refquote}

		\textbf{R:} We have removed much of the preferatory text explaining which quantity we are most interested in ($\lambda$) and now more explicitly state our goals at the end of the introduction. We hope that the transition to the methods is now less work for a reader.

	12. Ideas presented in L209, L262, L310 are important but explained too briefly [[done]]

@@

		\begin{refquote}
		L209ff This seems a really important set of ideas, but they are exemplified very briefly and very simplistically in the paper (L262, L310). I would have appreciated a much clearer explanation and example.
		\end{refquote}

		\textbf{R:} We are not quite sure what the Reviewer is asking here. $\alpha$ and $\beta$ truly are quite simple (as simple as the mean and standard deviation of a normal distribution) so there is not much more to say about them in themselves. We have expanded our discussion of different types of priors and how they may be obtained, and we hope that this clarifies matters for the Reviewer. We also note that a simple worked example, with accompanying R code, is provided in \emph{Appendix S3} so that readers can see exactly how to obtain $\alpha$ and $\beta$. In case the point of confusion is the idea that $\alpha$ and $\beta$ can be specified or derived depending on the type of prior desired, we have added a line (lines XX-XX) spelling this out.

		\begin{quotation}
			Parameters $\alpha$ and $\beta$ determine the shape of the prior distribution of $\lambda$, which follows a beta distribution. These are called hyper parameters and may be chosen to obtain an intentionally uninformative prior or derived using prior information to create an informed or empirical prior. 
			\end{quotation}

	13. Reframe line 241 in terms of solutions rather than difficulties [[done]]

@@

		\begin{refquote}
		L241 I wonder if you could phrase this more positively as ‘solutions’ rather than just ‘difficulties’?
		\end{refquote}

		\textbf{R:} We appreciate the Reviewer's desire for optimism, but we also selected a high-quality dataset specifically in order to demonstrate that empirically sampling all of the feasible interactions in a system is very unlikely. A lower-quality dataset would work just as well with our framework but would not highlight these difficulties so well. We have therefore kept this line as-is but have added another line (XX-XX) to show how the Bayesian framework provides a partial solution to the problems inherent in even high-quality empirical data.

		\begin{quotation}
		 The high spatiotemporal resolution of this dataset make it ideal for illustrating the difficulties in completely sampling a network; even with such an unusually high sampling effort, we anticipate that there will be many pairs of species which were rarely or never observed together. Using the Bayesian framework above, we can identify which potential interactions are more and less uncertain, allowing us to better predict the true structure of the metaweb.
		 \end{quotation}

	14. Line 283 seems to be a very important point and should be explained more clearly [[done]]

		\textbf{R:} Due to space constraints, we cannot provide an extensive description of Jensen's inequality here. As it is quite easy to look up the inequality by name, we are confident that interested readers can find a description to whatever level of detail suits their interest. 
		[[We could potentially add a demo in a new Appendix. Is this worthwhile?]]

### 2018-11-28 manuscript/CI_interactions.tex
most of reply done, now to look at structure

      We start from the perspective of a community ecologist faced with the task of describing a previously unknown interaction network. This ecologist will be interested in generating a description of the species/nodes present and the links between them~\citep{Roslin2016}.  Importantly, the information sought is conveyed by both the presence and \emph{absence} of links. Presences and absences are not, however, equally certain. An observed link definitely occurred, but there are multiple reasons why a given link may not be observed \emph{whether or not the interaction truly occurred}. In terms of the ones and zeros in a binary interaction matrix, we know that the ones are true but we cannot be sure that the zeros reflect interactions which actually do not occur. The detection of any interaction is a stochastic process. As a conceptual guide to the factors affecting this process, we describe three nested levels of uncertainty that roughly address the questions: "Could species $i$ and $j$ interact?", "Do they interact?" and "Do we observe the interaction?" (Fig.~\ref{conceptual_fig}).

@@

        The obvious rule of thumb to reduce uncertainty is ``sample more". Increasing sample sizes will reduce uncertainty about the upper bound of the interaction probability and will also increase the probability of detecting unlikely or cryptic interactions (e.g., interactions where $L$=1 but process or detection uncertainty is high). Despite these benefits, we note that there are limits to the utility of increased sampling. Since the probability of observing the co-occurrence of two species will always be higher than the probability of observing their interaction (since the probability of interaction is conditional on both interaction partners being present), we will accumulate observations of co-occurrences faster than we will accumulate observations of interactions (Fig.~\ref{histograms}C). Thus, the more we sample, the more zeros will appear in our interaction matrix.

        An added complication is that increased sampling will not reduce uncertainty evenly across interactions. To record an interaction between A and B, we need to identify both partners correctly (a non-trivial problem in many food webs; e.g.~\citealp{Kaartinen2011,Roslin2016}). For both molecular and rearing techniques, certain types of interactions may go unnoticed due to technical challenges~\citep{Wirta2014}. This can bias the set of recorded interactions towards those that are easier to observe~\citep{Carstensen2014,Jordano2016}.

        Most importantly, if two species are never observed co-occurring during several days of sampling then repeated sampling reveals nothing about their probability of interacting should they ever co-occur. Intensive sampling is, therefore, an important step towards reducing uncertainty in ecological networks but is unlikely to be enough in most cases. 
        Indeed, despite substantial improvements in recent decades, most available interaction networks are still undersampled and, moreover, are only subsets of larger communities~\citep{Jordano2016}.
        To supplement repeated sampling efforts and reduce uncertainty around unobserved interactions, we suggest researchers use prior knowledge of the same or similar systems. Here, we outline a Bayesian approach to combining repeated sampling and prior knowledge in probabilistic networks. We then demonstrate how to construct such a probabilistic network using an intensively and repeatedly sampled empirical dataset. Finally, we show how a probabilistic understanding of a network allows us to establish confidence intervals around measures of network structure.

@@

      Parameters $\alpha$ and $\beta$ determine the shape of the prior distribution of $\lambda$, which follows a beta distribution. These are called hyper parameters. The prior may be chosen to be uninformative, if no external information is available, or may draw on information such as distributions of connectance across webs, degree distributions, or trait-matching functions (see \emph{Appendix S3} for details and a simple quantitative example).

@@

        The relevant traits and degree of trait-matching between interaction partners is likely to vary between systems~\citep{Gravel2013,Weinstein2017a}. We therefore urge caution when using a model developed for one system as the prior for another.

@@

  \subsection*{An empirical example}
    [[Make the point of this clearer]]
      To illustrate the process of constructing a Bayesian network to quantify uncertainty about interactions, we use the comprehensively-sampled system of willows (\emph{Salix}), herbivorous gallers, and their natural enemies described by~\citet{Kopelke2017}. This dataset consists of a single community type sampled across Europe over 29 years and at 374 unique locations. The meta-network consists of 1,173 different interactions between 52 \emph{Salix} nodes, 92 herbivore nodes, and 126 natural enemy nodes (see \emph{Appendix S4} for details). 
      The high spatiotemporal resolution of this dataset make it ideal for illustrating the difficulties in completely sampling a network; even with such an unusually high sampling effort, we anticipate that there will be many pairs of species which were rarely or never observed together. To show the gaps in sampling, we compared the frequencies of observed co-occurrences and interactions. We then calculated an empirical prior and computed the posterior distribution of the probability of an as-yet-unobserved interaction being feasible ($\lambda$).
      We analysed both the \emph{Salix}-galler and galler-natural enemy components of the network but, for brevity, present only the latter here (see~\emph{Appendix S5} for \emph{Salix}-galler results).

@@

      Using the prior distributions and procedures described above, we calculated posterior probability distributions for species pairs that were not observed interacting. Using these posterior distributions and assuming probabilities of 1 for pairs of species that were observed interacting (see \emph{Appendix S2} for a justification), we created a suite of 100 webs by randomly sampling from each posterior distribution. After obtaining these posterior networks, we calculated the connectance of each web, as well as the mean number of links per galler and per natural enemy, and the nestedness (NODF) of the network.

>>>>

      We start from the perspective of a community ecologist faced with the task of describing a previously unknown interaction network. This ecologist will be interested in generating a description of the species/nodes present and the links between them~\citep{Roslin2016}.  Importantly, the information sought is conveyed by both the presence and \emph{absence} of links. Presences and absences are not, however, equally certain. An observed link definitely occurred, but there are multiple reasons why a given link may not be observed \emph{whether or not the interaction truly occurred}. Moreover, given an unobserved interaction in an empirical dataset, we often cannot determine why the interaction did not occur \emph{post hoc}. The detection of any interaction is a stochastic process subject to many levels of uncertainty. As a conceptual guide to the factors affecting this process, we describe three nested levels of uncertainty that roughly address the questions: "Could species $i$ and $j$ interact?", "Do they interact?" and "Do we observe the interaction?" (Fig.~\ref{conceptual_fig}).

@@

        [[move to box
        The obvious rule of thumb to reduce uncertainty is ``sample more". Increasing sample sizes will reduce uncertainty about the upper bound of the interaction probability and will also increase the probability of detecting unlikely or cryptic interactions (e.g., interactions where $L$=1 but process or detection uncertainty is high). Despite these benefits, we note that there are limits to the utility of increased sampling. Since the probability of observing the co-occurrence of two species will always be higher than the probability of observing their interaction (since the probability of interaction is conditional on both interaction partners being present), we will accumulate observations of co-occurrences faster than we will accumulate observations of interactions (Fig.~\ref{histograms}C). Thus, the more we sample, the more zeros will appear in our interaction matrix.

        An added complication is that increased sampling will not reduce uncertainty evenly across interactions. To record an interaction between A and B, we need to identify both partners correctly (a non-trivial problem in many food webs; e.g.~\citealp{Kaartinen2011,Roslin2016}). For both molecular and rearing techniques, certain types of interactions may go unnoticed due to technical challenges~\citep{Wirta2014}. This can bias the set of recorded interactions towards those that are easier to observe~\citep{Carstensen2014,Jordano2016}. [[Suggest combining approaches for better results - Jordano2016?]] ]]

        Most importantly, if two species are never observed co-occurring during several days of sampling then repeated sampling reveals nothing about their probability of interacting should they ever co-occur. Intensive sampling is, therefore, an important step towards reducing uncertainty in ecological networks but is unlikely to be enough in most cases. 
        Indeed, despite substantial improvements in recent decades, most available interaction networks are still undersampled and, moreover, are only subsets of larger communities~\citep{Jordano2016}.
        To supplement repeated sampling efforts and reduce uncertainty around unobserved interactions, we suggest researchers use prior knowledge of the same or similar systems. Here, we outline a Bayesian approach to combining repeated sampling and prior knowledge in order to assemble a metaweb. We then demonstrate this approach using an intensively and repeatedly sampled empirical dataset. Finally, we show how a probabilistic understanding of a metaweb allows us to establish confidence intervals around measures of network structure.

@@

      Parameters $\alpha$ and $\beta$ determine the shape of the prior distribution of $\lambda$, which follows a beta distribution. These are called hyper parameters and may be chosen to obtain an intentionally uninformative prior or derived using prior information to create an informed or empirical prior. The prior may be chosen to be uninformative, if no external information is available, or may draw on information such as distributions of connectance across webs, degree distributions, or trait-matching functions (see \emph{Appendix S3} for details and a simple quantitative example).

@@

        The relevant traits and degree of trait-matching between interaction partners is likely to vary between systems~\citep{Gravel2013,Weinstein2017a}. We therefore urge caution when using a model developed for one system as the prior for another. 
        See \emph{Appendix S3} for a simple quantitative example using degree distributions to obtain $\alpha$ and $\beta$.

@@

  \subsection*{An empirical example}
    [[Make the point of this clearer]]
      To illustrate the process of constructing a Bayesian network to quantify uncertainty about interactions, we use the comprehensively-sampled system of willows (\emph{Salix}), herbivorous gallers, and their natural enemies described by~\citet{Kopelke2017}. This dataset consists of a single community type sampled across Europe over 29 years and at 374 unique locations. The meta-network consists of 1,173 different interactions between 52 \emph{Salix} nodes, 92 herbivore nodes, and 126 natural enemy nodes (see \emph{Appendix S4} for details). 
      The high spatiotemporal resolution of this dataset make it ideal for illustrating the difficulties in completely sampling a network; even with such an unusually high sampling effort, we anticipate that there will be many pairs of species which were rarely or never observed together. Using the Bayesian framework above, we can identify which potential interactions are more and less uncertain, allowing us to better predict the true structure of the metaweb. To show the gaps in sampling, we compared the frequencies of observed co-occurrences and interactions. We then calculated an empirical prior and computed the posterior distribution of the probability of an as-yet-unobserved interaction being feasible ($\lambda$).
      We analysed both the \emph{Salix}-galler and galler-natural enemy components of the network but, for brevity, present only the latter here (see~\emph{Appendix S5} for \emph{Salix}-galler results).

@@

      Using the prior distributions and procedures described above, we calculated posterior probability distributions for species pairs that were not observed interacting. Using these posterior distributions and assuming probabilities of 1 for pairs of species that were observed interacting (see \emph{Appendix S2} for a justification), we created a suite of 100 webs by randomly sampling from each posterior distribution. Each of these webs is a prediction of the structure of the metaweb. After obtaining these posterior networks, we calculated the connectance of each web, as well as the mean number of links per galler and per natural enemy, and the nestedness (NODF) of the network.
